---
title: "Reproducible research peer assignment 1"
author: "Yasukazu IKAMI"
date: "Sunday, December 14, 2014"
output: html_document
---
# Reproducible research peer assignment 1

This is setting for assignment. 
```{r}
Sys.setlocale("LC_TIME", "English")
library(ggplot2)
```

## Loading and processing data
It is supposed that *activity.csv* is in the current directory.
```{r}
data <- read.csv("activity.csv")
```

Processing data to calculate total number of steps taken each day.
Dataframe *daytotal* consists col1 = date, col2 = total number of steps for the day.
```{r}
daytotal <- aggregate(steps ~ date, data, sum)
```

## What is mean total number of steps taken per day?
Making a histogram. x-axis is total number of steps taken per day.
```{r}
hist(daytotal$steps, xlab="Total number of steps taken per day",main="")
```

Calculation of mean of total number os steps taken per day.
```{r}
mean(daytotal$steps)
```

Calculation of median of total number of steps taken per day.
```{r}
median(daytotal$steps)
```

Results are very slightly different.

## What is the average daily activity pattern?
Processing data to calculate average activity pattern for all days.
Dataframe *intervalaverage* consists col1 = interval, col2 = average steps taken for all days.
```{r}
intervalaverage <- aggregate(steps ~ interval, data, mean)
```
Plot the time series of average number of steps taken, averaged across all days.
```{r}
plot(intervalaverage, type="l", ylab="Average number of steps",
     main="Average number of steps taken, averaged across all days")
```

Find the *interval* point where *steps* recorded their maximum on average.
```{r}
intervalaverage[which.max(intervalaverage$steps),"interval"]
```

## Imputing missing values
Calculate number of NA values from *steps* in dataframe *data*
```{r}
sum(is.na(data$steps)==TRUE)
```

Strategy for imputing missing values in original dataset is as follows. It is to replace NAs by average steps of the interval when NA value is recorded.
Processing data to calculate the average steps taken for each day. It is preparation for the inputing missing values.

Create temporary vector *filling*. It contains steps with NAs replaced with average steps for its interval.
When the element of *filling* is NA, first, find the interval of NA value is recorded by match() function.
And replace NA value by steps averaged by interval which is calculated former section.
```{r}
filling <- data$steps
for(i in 1:length(filling)){
    if(is.na(filling[i])){
        index <- match(data[i,"interval"],intervalaverage$interval)
        filling[i] <- intervalaverage[index,"steps"]
    }
}
```

Create new dataset *newdf* with replaced NAs. And rename colnames.
```{r}
newdf <- cbind(filling,data[2:3])
names(newdf) <- c("steps","date","interval")
```
Making a histogram of new dataset, calculating mean and median. These procedures are same as that for original dataset.
```{r}
newdaytotal <- aggregate(steps ~ date,newdf,sum)

hist(newdaytotal$steps,xlab="Total number of steps taken per day",main="")
mean(newdaytotal$steps)
median(newdaytotal$steps)
```
Mean and median value of total number of steps became the same by imputing NA values by the interval averaged steps.

## Are there differences in activity patterns between weekdays and weekends?
Dataframe *weekendlist* contains pairs of the day of the week and whether it is weekday or weekend.
```{r}
weekendlist <- data.frame(day=c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"),
                          we=c(rep("Weekday",5),rep("Weekend",2)))
```
New vector *wd* contains whether the day is weekday or weekend.
First, created as the copy of date of original dataset. And converted into the day (Sunday to Saturday).
```{r}
wd <- data$date
wd <- weekdays(as.Date(wd))
```
For all element of *wd*, find whether the element is weekday or weekend by matching *weekendlist*.
And replace the day by *weekday* or *weekend*.
```{r}
for(i in 1:length(wd)){
  index <- match(wd[i],weekendlist$day)
  wd[i] <- as.character(weekendlist[index,"we"])
  
}
wd <- as.factor(wd)
```
Final dataframe *n2df* is combination of NA replaced dataframe and the vector indicates the day is weekday or weekend.
*stepsintwd* is the dataframe contains col1=interval, col2 = weekday/weekend, col3 = average steps in each interval and weekday/weekend.
```{r}
n2df <- cbind(newdf,wd)
stepsintwd <- aggregate(steps ~ interval + wd,n2df,mean)
```

Two time series plots which show weekday and weekend average activity patterns averaged across all the days.
To plot panel plots, ggplot2 package is used.
```{r}
g <- ggplot(stepsintwd,aes(interval,steps))
g + geom_line() + facet_grid(wd ~ .) + labs(y="Average steps taken")
```

